Sparsity Weak Lensing Mass Reconstruction

### 2020-04-20 paper_ms_method_HSCY1/ms.tex
commi
make halolet
commit

\subsection{Density Contrast to Shear}
The lensing distortion $\kappa$ at a position $\vec{x}=(\vec{\theta}$, $\chi_s$) caused by the foreground inhomogeneous density distribution is 
\begin{equation}
\kappa(\vec{\theta},\chi_s)=\frac{3H_0^2\Omega_M}{2 c^2} \int_0^{\chi_s} d\chi_l \frac{\chi_l \chi_{sl}}{\chi_s} \frac{\delta(\chi_l \vec{\theta},\chi_l)}{a(\chi_l)}
\end{equation}
\citep{massMap-Glimpse3D2014}.
Substitute angular distance $D$ with redshift $z$ and we have
\begin{equation}\label{eq-delta2kappa}
\kappa(\vec{\theta},z_s)=\int_0^{z_s} dz_l \delta_{c}^{-1}(z_l,z_s)\delta(\vec{\theta},z_l).
\end{equation}
we term $\delta_{c}(z_l,z_s)$ as lensing critical density contrast and define it as
\begin{equation}
\delta_{c}^{-1}(z_l,z_s) =
\begin{cases}
\frac{3H_0\Omega_M}{2 c} \frac{\chi_l \chi_{sl} (1+z_l)}{ \chi_s E(z_l)} & (z_s>z_l),\\
0&(z_s \leq z_l).
\end{cases}
\end{equation}

@@

\subsection{Photo-$z$ Uncertainty}
Since the redshifts of source galaxies are estimated with a limited number of photometric bands, the estimated redshift of a galaxy suffer from large uncertainty. The probability distribution function (PDF) of a galaxies at position $(\vec{\theta},z_s)$ is denoted as $\mathcal{P}(z|z_s)$.  Note that, in order to simplify the future calculation, we assume the variation of the PDF across the transverse plane is small and neglect the dependency on the transverse position.

@@

\subsection{Dictionary}
Since many $N$-body simulations have shown the dark matter to be largely distributed in halos connected by filaments,
we assume that the density contrast field can be decomposed into multi-scaled NFW halo \citep{halo-NFW1997ApJ} and point mass at different positions
\begin{equation}
\delta(\vec{x}) = \sum_{s=0}^{N_s} \int d^3 x' \phi_s(\vec{x}-\vec{x}') \alpha_s(\vec{x'})\\
\end{equation}
where $\phi_0$ is $3$-D Dirac delta function representing point mass
\begin{equation}
\phi_0(\vec{x})= \delta_D(\theta_1) \delta_D(\theta_2) \delta_D(z).
\end{equation}
Since the scale of halo is much less than the reachable redshift resolution, we neglect the depth of halo on the line-of-sight direction as suggested by \citep{massMap-Glimpse3D2014}.
On the transverse plane, the NFW atom is modeled with the surface density profile of NFW halo with scale $\theta_s$ and truncation radius $c \theta_s$ \citep{haloModel-TJ2003-3pt}, where $c$ is generally known as concentration of NFW halo.  Whereas, the NFW atom is modeled with Dirac delta function on the line-of-sight direction
\begin{equation}
\begin{split}
\phi_s(\vec{x}) =&\frac{f }{2 \pi \theta_s^2 } F(|\vec{\theta}|/\theta_s) \delta_D(z),\\
&  (s=1..N_s)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{1}{(1-x^2)^{3/2}} \mathrm{arccosh} (\frac{x^2+c}{x(1+c)}) & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{1}{(x^2-1)^{3/2}} \mathrm{arccos} (\frac{x^2+c}{x(1+c)}) & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
and $f=1/[\ln (1+c)-c/(1+c)]$.

@@

\subsection{Loss Function with Constrains}
The loss function is defined as
\begin{equation}\label{eq-lossFun}
L = \norm{\gamma-\mathbf{M}\mathbf{P}\mathbf{Q}\mathbf{\Phi} \alpha}^2_2+\lambda \norm{\alpha}^1_1 + \tau\mathrm{TSV}(\alpha_0),
\end{equation}
where $\norm{\gamma-\mathbf{M}\mathbf{P}\mathbf{Q}\mathbf{\Phi} \alpha}^2_2$ is the normal chi-square, $\norm{\alpha^{N,P}}^1_1$ is the sparsity constrain, and $\mathrm{TSV}(\alpha^P)$ is the Total Squared Variation (TSV) constrain. $\norm{\bullet}_1$ and $\norm{\bullet}_2$ represent $l_1$ norm and $l_2$ norm, respectively.
The $\mathrm{TSV}$ operator is defined as 
\begin{equation}
\begin{split}
\mathrm{TSV}(\alpha_0) =&  \sum_{ijk} \{ (\alpha_0[i,j,k]-\alpha_0[i+1,j,k])^2\\
&+(\alpha_0[i,j,k]-\alpha_0[i,j+1,k])^2\\
&+(\alpha_0[i,j,k]-\alpha_0[i,j,k+1])^2 \},
\end{split}
\end{equation}
where $i=1..N_\theta$, $j=1..N_\theta$, are the pixel indexes for the two dimensions on the transverse plane, $k=1..N_l$ is the pixel index on the line-of-sight direction.

@@

\subsection{FISTA}
The basic FISTA algorithms is
\begin{equation}
X^{(n+1)}_m=\mathrm{ST}_{\lambda} (\alpha_m -\frac{\mu}{2A_{mm}} \frac{\partial f}{\partial \alpha_m})|_{\alpha^{(n)}_m} ,
\end{equation}
where 
\begin{equation}
\mathrm{ST}_{\lambda} (x) = \mathrm{sign} (x) \max (\abs{x}-\lambda,0)
\end{equation}
is the soft thresholding function.
\begin{equation}
\begin{split}
t^{(n+1)} &= \frac{1+\sqrt{1+4(t^{(n)})^2}}{2},\\
\alpha^{(n+1)}_m&=X^{(n+1)}_m+[\frac{t^{(n)}-1}{t^{(n+1)}}(X^{(n+1)}_m-X^{(n)}_m)],
\end{split}
\end{equation}
with $t^{(0)}=1$ and $X^{(0)}=\alpha^{(0)}$.
\section{Simulation}

\section{Results}
...
\section{Summary}
...
\bibliographystyle{aasjournal}
\bibliography{weak_lensing,lensCat,cosmoSim,cmbObs,gglens,shearCor,massMap,other}

>>>>

\subsection{Density Contrast to Shear}
The lensing distortion $\kappa$ at a position $\vec{x}=(\vec{\theta}$, $\chi_s$) caused by the foreground inhomogeneous density distribution is

\begin{figure*}
 \includegraphics[width=0.95\textwidth]{nfwlet-atom-2D.pdf}
 \caption{The NFW atom with different scale radius ($r_s$). The first row shows the NFW atoms in Fourier space and the second row shows the NFW atom in Real space.}
\end{figure*}

\begin{equation}
\kappa(\vec{\theta},\chi_s)=\frac{3H_0^2\Omega_M}{2 c^2} \int_0^{\chi_s} d\chi_l \frac{\chi_l \chi_{sl}}{\chi_s} \frac{\delta(\chi_l \vec{\theta},\chi_l)}{a(\chi_l)}
\end{equation}
\citep{massMap-Glimpse3D2014}.
Substitute angular distance $D$ with redshift $z$ and we have
\begin{equation}\label{eq-delta2kappa}
\kappa(\vec{\theta},z_s)=\int_0^{z_s} dz_l \delta_{c}^{-1}(z_l,z_s)\delta(\vec{\theta},z_l).
\end{equation}
we term $\delta_{c}(z_l,z_s)$ as lensing critical density contrast and define it as
\begin{equation}
\delta_{c}^{-1}(z_l,z_s) =
\begin{cases}
\frac{3H_0\Omega_M}{2 c} \frac{\chi_l \chi_{sl} (1+z_l)}{\chi_{s} E\left(z_l\right)} & (z_s>z_l),\\
0&(z_s \leq z_l).
\end{cases}
\end{equation}

@@

\begin{figure}
 \includegraphics[width=0.5\textwidth]{nfwlet-atom-1D.pdf}
 \caption{The $1$-D slices of NFW atoms with different scale radius ($r_s=0,1,2,4,8$).}
\end{figure}
\subsection{Photo-$z$ Uncertainty}
Since the redshifts of source galaxies are estimated with a limited number of photometric bands, the estimated redshift of a galaxy suffer from large uncertainty. The probability distribution function (PDF) of a galaxies at position $(\vec{\theta},z_s)$ is denoted as $\mathcal{P}(z|z_s)$.  Note that, in order to simplify the future calculation, we assume the variation of the PDF across the transverse plane is small and neglect the dependency on the transverse position.

@@

\subsection{Dictionary}
Since many $N$-body simulations have shown the dark matter to be largely distributed in halos connected by filaments,
we assume that the density contrast field can be decomposed into multi-scaled NFW halo \citep{halo-NFW1997ApJ} and point mass at different positions
\begin{equation}
\delta(\vec{x}) = \sum_{s=0}^{N_s} \int d^3 x' \phi_s(\vec{x}-\vec{x}') \alpha_s(\vec{x'})\\
\end{equation}
where $\phi_0$ is $3$-D Dirac delta function representing point mass
\begin{equation}
\phi_0(\vec{x})= \delta_D(\theta_1) \delta_D(\theta_2) \delta_D(z).
\end{equation}
Since the scale of halo is much less than the reachable redshift resolution, we neglect the depth of halo on the line-of-sight direction as suggested by \citep{massMap-Glimpse3D2014}.
On the transverse plane, the NFW atom is modeled with the surface density profile of NFW halo with scale $\theta_s$ and truncation radius $c \theta_s$ \citep{haloModel-TJ2003-3pt}, where $c$ is generally known as concentration of NFW halo.  Whereas, the NFW atom is modeled with Dirac delta function on the line-of-sight direction

\begin{equation}
\begin{split}
\phi_s(\vec{x}) =&\frac{f }{2 \pi \theta_s^2 } F(|\vec{\theta}|/\theta_s) \delta_D(z),\\
&  (s=1..N_s)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{1}{(1-x^2)^{3/2}} \mathrm{arccosh} (\frac{x^2+c}{x(1+c)}) & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{1}{(x^2-1)^{3/2}} \mathrm{arccos} (\frac{x^2+c}{x(1+c)}) & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
and $f=1/[\ln (1+c)-c/(1+c)]$.

@@

\subsection{Loss Function with Constrains}
\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their mlz best photo-$z$ estimation. The blue histogram is the number distribution of the best photo-$z$ estimation. The vertical dashed lines are the bounds of bins. The galaxies are evenly distributed in each bins.}
\end{figure}
The loss function is defined as
\begin{equation}\label{eq-lossFun}
L = \norm{\gamma-\mathbf{M}\mathbf{P}\mathbf{Q}\mathbf{\Phi} \alpha}^2_2+\lambda \norm{\alpha}^1_1 + \tau\mathrm{TSV}(\alpha_0),
\end{equation}
where $\norm{\gamma-\mathbf{M}\mathbf{P}\mathbf{Q}\mathbf{\Phi} \alpha}^2_2$ is the normal chi-square, $\norm{\alpha^{N,P}}^1_1$ is the sparsity constrain, and $\mathrm{TSV}(\alpha^P)$ is the Total Squared Variation (TSV) constrain. $\norm{\bullet}_1$ and $\norm{\bullet}_2$ represent $l_1$ norm and $l_2$ norm, respectively.
The $\mathrm{TSV}$ operator is defined as
\begin{equation}
\begin{split}
\mathrm{TSV}(\alpha_0) =&  \sum_{ijk} \{ (\alpha_0[i,j,k]-\alpha_0[i+1,j,k])^2\\
&+(\alpha_0[i,j,k]-\alpha_0[i,j+1,k])^2\\
&+(\alpha_0[i,j,k]-\alpha_0[i,j,k+1])^2 \},
\end{split}
\end{equation}
where $i=1..N_\theta$, $j=1..N_\theta$, are the pixel indexes for the two dimensions on the transverse plane, $k=1..N_l$ is the pixel index on the line-of-sight direction.

@@

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{mlz-poz.pdf}
 \caption{The PDF of photo-$z$ error for $10$ source redshift bins.}
\end{figure}

\begin{figure}
 \centering
 \includegraphics[width=0.5\textwidth]{lensing_efficiency.pdf}
 \caption{The blue line shows the normalized lensing efficiency for lens bin at $z_{l}=0.51$. The orange line is the normalized lensing efficiency while taking into account the photo-$z$ uncertainty.}
\end{figure}
\subsection{FISTA}
The basic FISTA algorithms is
\begin{equation}
X^{(n+1)}_m=\mathrm{ST}_{\lambda} (\alpha_m -\frac{\mu}{2A_{mm}} \frac{\partial f}{\partial \alpha_m})|_{\alpha^{(n)}_m} ,
\end{equation}
where
\begin{equation}
\mathrm{ST}_{\lambda} (x) = \mathrm{sign} (x) \max (\abs{x}-\lambda,0)
\end{equation}
is the soft thresholding function.
\begin{equation}
\begin{split}
t^{(n+1)} &= \frac{1+\sqrt{1+4(t^{(n)})^2}}{2},\\
\alpha^{(n+1)}_m&=X^{(n+1)}_m+[\frac{t^{(n)}-1}{t^{(n+1)}}(X^{(n+1)}_m-X^{(n)}_m)],
\end{split}
\end{equation}
with $t^{(0)}=1$ and $X^{(0)}=\alpha^{(0)}$.

\section{HSC-like Simulation}

\subsection{Shear Measurement Error}
\subsection{Photo-$z$ Error}

\section{Results}

\section{Summary}

\bibliographystyle{aasjournal}
\bibliography{weak_lensing,lensCat,cosmoSim,cmbObs,gglens,shearCor,massMap,other}
